\section{From groups to groupoids}

\subsection{Motivation}

One possible limitation coming from searching for Cayley subgraphs is that they are order-regular \ie the in- and the out-degree $d = |\cu|$ of each vertex is the same. That is, for a general graph $G$, the size of the weight kernel $w$ of an (EC*) convolution operator $f_w$ supported on $\cu$ is bounded by $d$, which in turn is bounded by twice the mininimal degree of $G$ (twice because $G$ is undirected and $\cu$ can contain every inverse).

There are a lot of possible strategies to overcome this limitation. For example:
\begin{enumerate}
  \item connecting each vertex with its $k$-hop neighbors, with $k > 1$,
  \item artificially creating new connections for less connected vertices,
  \item allowing the supporting set $\cn$ to exceed $\cu$ \ie dropping * in (EC*).
\end{enumerate}

These strategies require to concede that the topological structure supported by $G$ is not the best one to support an (EC*) convolution on it, which breeds the following question:
\begin{itemize}
  \item What can we relax in the previous (EC*) contruction in order to unbound the supporting set, and still preserve the equivariance characterization?
\end{itemize}

The latter constraint is a consequence that every vertex of the Cayley subgraph $\vec{G}$ must be composable with every generator from $\cu$. Therefore, an answer consists in considering groupoids~\citep{brandt1927verallgemeinerung} instead of groups. Roughly speaking, a groupoid is almost a group except that its composition law needs not be defined everywhere. \cite{weinstein1996groupoids}, unveiled the benefits to base convolutions on groupoids instead of groups in order to exploit partial symmetries.

\subsection{Definition of notions related to groupoids}

\begin{definition}\textbf{Groupoid}\\
A \emph{groupoid} $\Upsilon$ is a set equipped with a partial composition law with domain $\cd \subset \Upsilon \times \Upsilon$, called \emph{composition rule}, that is
\begin{enumerate}
  \item\label{enum:g1} closed into $\Upsilon$ \ie $\forall (g, h) \in \cd, gh \in \Upsilon$
  \item\label{enum:g2} associative \ie
    $\forall f,g,h \in \Upsilon,
      \begin{cases}
        (f, g), (g, h) \in \cd \Leftrightarrow (fg,h), (f, gh) \in \cd\\
        (f, g), (fg, h) \in \cd \Leftrightarrow (g,h), (f, gh) \in \cd\\
        \text{when defined, } (fg)h = f(gh)
      \end{cases}$ \label{enum:2}
  \item\label{enum:g3} invertible \ie
    $\forall g \in \Upsilon, \exists! g^{-1} \in \Upsilon \quad\st
      \begin{cases}
        (g,g^{-1}), (g^{-1},g)  \in \cd\\
        (g,h) \in \cd \Rightarrow g^{-1}gh = h\\
        (h,g) \in \cd \Rightarrow hgg^{-1} = h
      \end{cases}$
\end{enumerate}
Optionally, it can be \emph{domain-symmetric} \ie $(g,h) \in \cd \Leftrightarrow (h,g) \in \cd$, and \emph{abelian} \ie domain-symmetric with $gh = hg$.
\end{definition}

\begin{remark}
Note that left and right inverses are necessarily equal (because $(gg^{-1})g=g(g^{-1}g)$). Also note we can define a right identity element $e^r_g = g^{-1}g$, and a left one $e^l_g = gg^{-1}$, but they are not necessarily equal and depend on $g$.
\end{remark}

Most definitions related to groups can be adapted to groupoids. In particular, let's adapt a few notions.

\begin{definition}\textbf{Groupoid partial action}\\
A partial \emph{action} of a groupoid $\Upsilon$ on a set $V$, is a function $L$, with domain $\cd_L \subset \Upsilon \times V$ and valued in $V$, such that the map $g \mapsto L_g$ is a groupoid homomorphism.
\end{definition}

\begin{remark}
As usual, we will confound $L_g$ and $g$ when there is no possible confusion, and we denote $\cd_{L_g} = \cd_g = \{ v \in V, (g,v) \in \cd_L \}$.
\end{remark}

\begin{definition}\textbf{Partial equivariant map}\\
A map $\varphi$ from a groupoid $\Upsilon$ partially acting on the destination set $V$ is said to be a \emph{partial equivariant map} if
\begin{gather*}
\forall g,h \in \Upsilon,
  \begin{cases}
    \varphi(h) \in \cd_g \Leftrightarrow (g,h) \in \cd\\
    g(\varphi(h)) = \varphi(gh)
  \end{cases}
\end{gather*}
\end{definition}

Also, $\varphi$-equivalence between a subgroupoid and a set is defined similarly with $\varphi$ being a bijective \emph{partial} equivariant map between them.

\begin{definition}\textbf{Partial transformations groupoid}\\
The \emph{partial transformations groupoid} $\Psi^*(V)$, is the set of invertible partial transformations, equipped with the functional composition law with domain~$\cd$ such that
\begin{gather*}
\begin{cases}
\cd_{gh} = h(\cd_h) \cap \cd_g\\
(g,h) \in \cd \Leftrightarrow \cd_{gh} \neq \emptyset
\end{cases}
\end{gather*}
\end{definition}

\begin{remark}Note that a subgroupoid $\Upsilon \subset \Psi^*(V)$ is domain-symmetric when $\exists v \in V, g(v) \in \cd_h \Leftrightarrow \exists u \in V, h(u) \in \cd_g$
\end{remark}

\subsection{Construction of partial convolutions}

The expression of the convolution we constructed in the previous section cannot be applied as is. We first need to extend the algebraic objects we work with. 
Extending a partial transformation $g$ on the signal space $\cs(V)$ (and thus the convolutions) is a bit tricky, because only the signal entries corresponding to $\cd_g$ are moved. A convenient way to do this is to consider the groupoid closure obtained with the addition of an absorbing element.

\begin{definition}\textbf{Zero-closure}\\
The \emph{zero-closure} of a groupoid $\Upsilon$, denoted $\Upsilon^0$, is the set $\Upsilon \cup {0}$, such that the groupoid axioms \ref{enum:g1}, \ref{enum:g2} and \ref{enum:g3}, and the domain $\cd$ are left unchanged, and
\begin{enumerate}\setcounter{enumi}{3}
\item the composition law is extended to $\Upsilon^0 \times \Upsilon^0$ with $\forall (g,h) \notin \cd, gh = 0$
\end{enumerate}
\end{definition}

\begin{remark}Note that this is coherent as the properties \ref{enum:g2} and \ref{enum:g3} are still partially defined on the original domain $\cd$.
\end{remark}

Now, we will also extend every other algebraic object used in the expression of the $\varphi$-convolution and the $\M$-convolution, so that we can directly apply our previous constructions.

\begin{lemma}\textbf{Extension of $\varphi$ on $V^0$}\\
Let a partial equivariant map $\varphi: \Upsilon \rightarrow V$. It can be extended to a (total) equivariant map $\varphi: \Upsilon^0 \rightarrow V^0 = V \cup \varphi(0)$, such that $\varphi(0) \notin V$, that we denote $0_V = \varphi(0)$, and such that
\begin{gather*}
	\forall g \in \Upsilon^0, \forall v \in V^0, g(v) = \begin{cases} \varphi(gg_v) &\text{ if } g_v \in \cd_g\\ 0_V &\text{ else}\end{cases}
\end{gather*}
\end{lemma}
\begin{proof}
We have $\varphi(0) \notin V$ because $\varphi$ is bijective. Additionaly, we must have $\forall (g,h) \notin \cd, g(\varphi(h)) = \varphi(gh) = \varphi(0) = 0_V$.
\end{proof}

\begin{remark} Note that for notational conveniency, we may use the same symbol~$0$ for~$0_{\Upsilon}$, $0_V$ and~$0_{\bbr}$.
\end{remark}

Similarly to $\Phi^*(V)$, $\Psi^*(V)$ can also move signals of $\cs(V)$.

\begin{lemma}\textbf{Extension of injective partial transformations to $\cs(V)$}\\
Let $g \in \Psi^*(V)$. Its extension is done in two steps:
\begin{enumerate}
  \item $g$ is extended to $V^0 = V \cup \{0_V\}$ as $g(v) = 0_V \Leftrightarrow v \notin \cd_g$.
  \item Under the convention $\forall s \in \cs(V), s[0_V] = 0_\bbr$, $g$ is extended via linear extension to $\cs(V)$, and we have
  \begin{gather*}
  \forall s \in \cs(V), \forall v \in V, g(s)[v] = s[g^{-1}(v)]
  \end{gather*}
\end{enumerate}
\end{lemma}
\begin{proof}
Straightforward.
\end{proof}

With these extensions, we can obtain the partial $\varphi$- and $\M$-convolutions related to~$\Upsilon$ almost by substituting~$\Upsilon^0$ to~$\Gamma$ in \defref{def:convv} and \defref{def:convm}.

\begin{definition}\textbf{Partial convolution}\\
Let a subgroupoid $\Upsilon \subset \Psi^*(V)$, such that $\Upsilon \overset\varphi\equiv V$. The partial $\varphi$- and $\M$-convolutions, based on~$\Upsilon$, are defined on its zero-closure, with the same expression as if~$\Upsilon^0$ were a subgroup, and by extension of~$\varphi$ and of the groupoid partial actions \ie
\begin{enumerate}[label=(\roman*)]
  \item $\forall s, w \in \cs(V),
  			s \ast_\varphi w = \displaystyle\sum_{v \in V} s[v] \h{2} g_v(w) = \displaystyle\sum_{g \in \Upsilon} s[\varphi(g)] \h{2} g(w)$ \label{enum:i}
  \item $\forall (w,s) \in \cs(\Upsilon) \times \cs(V), w \ast_{\M} s  = \displaystyle\sum_{g \in \Upsilon} w[g] \h{2} g(s)$ \label{enum:ii}
\end{enumerate}
\end{definition}

\paragraph{Symmetrical expressions}
Note that, as $\forall r, r[0] = 0$, the partial convolutions can also be expressed on the domain~$\cd$ with a convenient symmetrical expression:
\begin{enumerate}[label=(\roman*)]
  \item $\forall u \in V, (s \ast_\varphi w)[u] =  \displaystyle\sum_{\substack{(g_a,g_b) \in \cd\\ \st g_ag_b = g_u}} s[a] \h{2} w[b]$
  \item $\forall u \in V, (w \ast_{\M} s)[u]  = \displaystyle\sum_{\substack{v \in \cd_g\\ \st g(v) = u}} w[g] \h{2} s[v]$
\end{enumerate}

We obtain an equivariance characterization similar to \propref{prop:equiG} and \corref{cor:equiM}.

\begin{proposition}\textbf{Characterization by equivariance to $\Upsilon$}\\
Let a subgroupoid $\Upsilon \subset \Psi^*(V)$, such that~$\Upsilon \overset\varphi\equiv V$, with $\ast$ based on~$\Upsilon$.
\begin{enumerate}
\item Then,
\begin{enumerate}[label=(\roman*)]
  \item partial $\varphi$-convolution right-operators are equivariant to~$\Upsilon$,\label{enum:i}
  \item if $\Upsilon$ is abelian, partial $\M$-convolution left-operators are equiv to~$\Upsilon$.\label{enum:ii}
\end{enumerate}
\item Conversely, \label{enum:2}
\begin{enumerate}[label=(\roman*)]
  \item if $\Upsilon$ is domain-symmetric, linear transformations of $\cs(V)$ that are equivariant to~$\Upsilon$ are partial $\varphi$-convolution right-operators,
  \item if $\Upsilon$ is abelian, they are also partial $\M$-convolution left-operators.
\end{enumerate}
\end{enumerate}
\label{prop:equiP}
\end{proposition}
\begin{proof}
\begin{enumerate}[label=(\roman*)]
   \item
  \begin{enumerate}
    \item Direct sense:\\
    Using the symmetrical expressions, and the fact that $\forall r, r[0] = 0$, we have
    \begin{align*}
    (f_w \circ g (s))[u] & = \displaystyle\sum_{\substack{(g_a,g_b) \in \cd\\ \st g_ag_b = g_u}} g(s)[a] \h{2} w[b]\\
    				& = \displaystyle\sum_{\substack{(g_a,g_b) \in \cd\\ \st g_ag_b = g_u}} s[g^{-1}(a)] \h{2} w[b]\\
    				& = \displaystyle\sum_{\substack{(g_a,g_b) \in \cd\\ \st (g, g_a) \in \cd\\ \st gg_ag_b = g_u}} s[a] \h{2} w[b]\\
    				& = \displaystyle\sum_{\substack{(g_a,g_b) \in \cd\\ \st (g, g_a) \in \cd\\ \st g_ag_b = g^{-1}g_u = g_{\varphi(g^{-1}g_u)} = g_{g^{-1}(u)} }} s[a] \h{2} w[b]\\
    				& = f_w (s)[g^{-1}(u)]\\
    				& = (g \circ f_w (s))[u]
    \end{align*}

    \item Converse:\\
    Let $v \in V$. Denote $e^r_{g_v} = g_v^{-1}g_v$ the right identity element of $g_v$, and $e^r_v = \varphi(e^r_{g_v})$. We have that
    \begin{gather*}
    g_v(e^r_v) = v\\
    \text{So, } \delta_v = g_v(\delta_{e^r_v})
    \end{gather*}

    Let $f \in \cl(\cs(V))$ that is equivariant to $\Upsilon$, and $s \in \cs(V)$. Thanks to the previous remark we obtain that
    \begin{align}
    f(s) & = \displaystyle\sum_{v \in V} s[v] \h{2} f(\delta_v)\nonumber\\
         & = \displaystyle\sum_{v \in V} s[v] \h{2} f(g_v(\delta_{e^r_v}))\nonumber\\
         & = \displaystyle\sum_{v \in V} s[v] \h{2} g_v(f(\delta_{e^r_v}))\nonumber\\
         & = \displaystyle\sum_{v \in V} s[v] \h{2} g_v(w_v) \label{eq:last}
    \end{align}
    where $w_v = f(\delta_{e^r_v})$. In order to finish the proof, we need to find $w$ such that $\forall v \in V, g_v(w) = g_v(w_v)$.

    Let's consider the equivalence relation $\ccr$ defined on $V \times V$ such that:
    \begin{align}
    a \ccr b & \Leftrightarrow    w_a = w_b\nonumber\\
            & \Leftrightarrow e^r_a = e^r_b\nonumber\\
            & \Leftrightarrow g_a^{-1}g_a = g_b^{-1}g_b\nonumber\\
            & \Leftrightarrow (g_b, g_a^{-1}) \in \cd\nonumber\\
            & \Leftrightarrow (g_a^{-1}, g_b) \in \cd \label{eq:eq}
    \end{align}
    with \eqref{eq:eq} owing to the fact that $\Upsilon$ is domain-symmetric.

    Given $x\in V$, denote its equivalence class $\ccr(x)$. Under the hypothesis of the axiom of choice~\citep{zermelo1904beweis} (if $V$ is infinite), define the set $\aleph$ that contains exactly one representative per equivalence class. Let $w = \sum_{n \in \aleph} w_n$. Then $V$ is the disjoint union $V = \displaystyle\cup_{n \in \aleph} \ccr(n)$ and \eqref{eq:last} rewrites:
    \begin{align}
    \forall u \in V, f(s)[u] & = \displaystyle\sum_{n \in \aleph}\sum_{v \in \ccr(n)} s[v] \h{2} g_v(w_n)[u]\nonumber\\
    	 				  & = \displaystyle\sum_{n \in \aleph}\sum_{v \in \ccr(n)} s[v] \h{2} w_n[g_v^{-1}(u)]\nonumber\\
    	 				  & = \displaystyle\sum_{n \in \aleph}\sum_{v \in \ccr(n)} s[v] \h{2} w[g_v^{-1}(u)]\label{eq:eqq}\\
    	 				  & = (s \ast_\varphi w)[u]\nonumber
    \end{align}
    where \eqref{eq:eqq} is obtained thanks to \eqref{eq:eq}.
  \end{enumerate}

  \item With symmetrical expressions, it is clear that the convolution is abelian, if and only if, $\Upsilon$ is abelian. Then \ref{enum:i} concludes.
\end{enumerate}
\end{proof}

\paragraph{Inclusion of (EC)}
Similarly to the construction in \secref{sec:edges}, partial convolutions can define (EC) and (EC*) counterparts with a characterization of admissibility by groupoid Cayley subgraph isomorphism, and similar intrinsic properties.

\paragraph{Limitation of partial convolutions}
However, because of the groupoid associativity, if $g \in \Psi_{\EC}^*(G)$, then, any $v \in V$ \st $g(u) = v$ would be constrained to allow to be acted by every $h$ \st $(h,g) \in \cd$, which fails at unbounding the supporting set of a partial (EC*)~convolutions.

\subsection{Construction of path convolutions}
\label{sec:path}

To answer the limitation of partial convolutions, given $g \in \langle \cu \rangle$ where $\cu \subset \Psi_{\EC}^*(G)$, the idea is to proceed with a foliation of $g$ into pieces, each corresponding to an edge $e \in E$, and together generating another groupoid with a different associativity law, as follows.

\begin{definition}\textbf{Path groupoid}\\
Let $\cu \subset \Psi_{\EC}^*(G)$. The \emph{path groupoid} generated from $\cu$, denoted $\cu \ltimes G$, with composition rule $\cdl$, is the groupoid obtained inductively as:
\begin{enumerate}
  \item $\cu \ltimes_1 G = \{(g,v) \in \cu \times V, v \in \cd_g \} \subset \cu \ltimes G$
  \item $((g_n,v_n) \cdots (g_1,v_1) , (h_m,u_m) \cdots (h_1,u_1)) \in \cdl \Leftrightarrow h_m(u_m) = v_1$
  \item $(g_n,v_n) \cdots (g_1,v_1) \in \cu \ltimes G \Rightarrow (g_1^{-1}, g_1(v_1)) \cdots (g_n^{-1}, g_n(v_n)) \in \cu \ltimes G$
\end{enumerate}
Call path its objects. Given a length $l \in \bbn^*$, denote $\cu \ltimes_l G$ the subset composed of the paths that are the composition of exactly~$l$ paths of $\cu \ltimes_1 G$.
\end{definition}

\begin{remark}
This groupoid construction is inspired from the field of operator algebra where partial action groupoids have been extensively studied, \eg \cite{nica1994groupoid,exel1998partial,li2016partial}.
\end{remark}

Such groupoids usually come equipped with source and target maps. We also define the path map.

\begin{definition}\textbf{Source, target and path maps}\\
Let a path groupoid $\cu \ltimes G$. We define on it the \emph{source map}~$\alpha$ the \emph{target map}~$\beta$ and the \emph{path map}~$\gamma$ as:
\begin{gather*}
\begin{cases}
  \alpha: (g_n,v_n) \cdots (g_1,v_1) \mapsto v_1 \in V\\
  \beta: (g_n,v_n) \cdots (g_1,v_1) \mapsto g_n(v_n) \in V\\
  \gamma: (g_n,v_n) \cdots (g_1,v_1) \mapsto g_ng_{n-1}\ldots g_1 \in \Psi^*(V^0)\\
  %\lambda: (g_1,v_1) \cdots (g_n,v_n) \mapsto n \in \bbn^*
\end{cases}
\end{gather*}
\end{definition}

\begin{remark}Note that the path groupoid can also be obtained by derivation of the partial transformation groupoid (\ie $p \in \cu \ltimes G$ can be seen as a derivative of $\gamma(p)$ \wrt $\alpha(p)$), and can thus be seen as the local structure of it.
\end{remark}

\begin{lemma}\h{0}\\
Note the following properties:
\begin{enumerate}
  \item $(p,q) \in \cdl \Leftrightarrow \beta(p) = \alpha(q)$
  \item $\alpha(p) = \beta(p^{-1})$
  \item $\gamma$ is a groupoid partial action. We will denote $\gamma_p$ instead of $\gamma(p)$. \label{enum:3}
  %\item $\forall (p,q) \in \cdl, \gamma_{pq} = \gamma_{p}\gamma_{q}$
  %\item $\beta$ is a partial equivariant map for the groupoid partial action $\gamma$ on $V$.
\end{enumerate}
\end{lemma}

\begin{remark}
Note that this time we won't use the notation $p(v)$ for $\gamma_p(v)$ in order to better differentiate between the composition laws in $\langle \cu \rangle$ and $\cu \ltimes G$.
\end{remark}

One of the key object of our contruction is the use of $\varphi$-equivalence in order to transform a sum over a group(oid) of (partial) transformations, into a sum over the vertex set. With the current notion of path groupoid, searching for something similar amounts to searching for a graph traversal.

\begin{definition}\textbf{Traversal set}\\
Let a graph $\gve$ that is connected. A \emph{traversal set} is a pair $(\cu, \ct)$ of (EC) partial transformations subsets $\subset \Psi_{\EC}^*(G)$, such that
\begin{enumerate}
  \item An edge can only correspond to a unique $g\in \cu$,\\
    \ie $\forall g,h \in \cu : \exists v \in V, g(v) = h(v) \Rightarrow g=h$
  \item The (EC) partial transformations of $\ct$ are restrictions of those of $\cu$,\\
    \ie $\forall g \in \cu, \exists! h \in \ct, \begin{cases}\cd_h \subset \cd_g\\ \forall v \in \cd_h, h(v) = g(v)\end{cases}$, \\
    (equivalently, $\ct \ltimes G$ is a subgroupoid of $\cu \ltimes G$)
  % \item The edges corresponding to $\ct$ are included in those of $\cu$,\\
  %   \ie $\ct \ltimes_1 G \subset \cu \ltimes_1 G$
  \item The subgraph $G_{\ct} = \langle V, \ct \ltimes_1 G \rangle$ is a covering tree of $G$.
\end{enumerate}
We denote $(\cu, \ct) \in \tree(G)$, and denote by $r$ the root of $G_{\ct}$.
\end{definition}

\begin{remark}The assumption that the graph $G$ is connected has been made. This doesn't lose generality as the construction can be replicated to each connected component in the general case.
\end{remark}

A traversal set $(\cu, \ct)$ defines a $\varphi$-equivalence between the $\alpha$-fiber of the root~$r$ and the vertex set~$V$ as follows.

\begin{lemma}\textbf{Path $\varphi$-Equivalence}\\
Let $(\cu, \ct) \in \tree(G)$. Given $v \in V$, there exists a unique $p_v \in \ct \ltimes G$ such that $\alpha(p_v) = r$ and $\beta(p_v) = v$. Define $\varphi: p_v \mapsto v$. Then $\varphi: \alpha^{-1}_{\ct \ltimes G}\{r\} \rightarrow V$ is a bijective partial equivariant map.
\end{lemma}
\begin{proof}
Bijectivity is a consequence of the covering tree structure of $\ct$.
Equivariance because $\gamma_{p_v}(u) = \gamma_{p_v}\gamma_{p_u}(r) = \gamma_{p_vp_u}(r) = \varphi(p_vp_u)$.
\end{proof}

We can now define the convolution that is based on a path groupoid.

\begin{definition}\textbf{Path convolution}\\
Let $(\cu, \ct) \in \tree(G)$. The \emph{path convolution} is the partial convolution based on the path subgroupoid $\ct \ltimes G$, which uses the groupoid partial action $\gamma := \gamma^{\cu \ltimes G}$ of the embedding groupoid $\cu \ltimes G$.
\begin{enumerate}[label=(\roman*)]
  \item In what follows are the three expressions of the path $\varphi$-convolution for signals $s_1, s_2 \in \cs(V)$, and $u \in V$:
\begin{align*}
(s \ast_\varphi w) & = \displaystyle\sum_{v \in V} s[v] \h{2} \gamma_{p_v}(w)\\
                   & = \displaystyle\sum_{\substack{p \in \ct \ltimes G\\ \st \alpha(p) = r}} s[\varphi(p)] \h{2} \gamma_{p}(w)\\
(s \ast_\varphi w)[u] & = \displaystyle\sum_{\substack{(a,b) \in V\\ \st \gamma_{p_a}(b)=u}} s[a] \h{2} w[b]
\end{align*}
  \item The mixed formulations with $w \in \cs(\ct \ltimes G)$ are:
\begin{align*}
(w \ast_{\M} s) & = \displaystyle\sum_{\substack{p \in \ct \ltimes G\\ \st \alpha(p) = r}} w[p] \h{2} \gamma_{p}(s)\\
(w \ast_{\M} s)[u] & = \displaystyle\sum_{\substack{(p,v) \in \ct \ltimes G \times V\\ \st \alpha(p) = r\\ \st \gamma_{p}(v)=u}} w[p] \h{2} s[v]
\end{align*}
\end{enumerate}
\end{definition}

\begin{remark}The role of $\ct$ is to provide a $\varphi$-equivalence. The role of $\cu$ is to extend every partial transformation $\gamma^{\ct \ltimes G}_g$ to the domain of its unrestricted counterpart $\gamma^{\cu \ltimes G}_g$.
\end{remark}

\propref{prop:equiP} also holds for path groupoids, except that the domain-symmetric condition of 2.(i) is not needed.

\begin{proposition}\textbf{Characterization by equivariance to $\cu \ltimes G$'s action}\\
Let $(\cu, \ct) \in \tree(G)$.
\begin{enumerate}[label=(\roman*)]
\item The class of linear transformations of $\cs(V)$ that are equivariant to the path actions of $\cu \ltimes G$ is exactly the path $\varphi$-convolution right-operators;
\item in the abelian case, they are also exactly the $\M$-convolution left-operators.
\end{enumerate}
\end{proposition}

\begin{proof}
Instead of the domain-symmetric condition that was used in the proof of the converse of \propref{prop:equiP} (2.(i)), we use the fact that any vertex can be reached with an action from the root of the covering tree of the traversal set. Indeed, given $v \in V$, as we have $\gamma_{p_v}(r)=v$, then $\gamma_{p_v}(\delta_r) = \delta_v$. Therefore, by developping a linear transformation $f(s)$ on the dirac family, and commuting $f$ with $\gamma_{p_v}$, we obtain that $f(s) = s \ast_\varphi w$, where $w = f(\delta_r)$. The rest of the proof is similar to that of \propref{prop:equiP}.
\end{proof}

\begin{remark}
Note that $\cu \ltimes V$'s action is almost the same as the groupoid partial action of $\Upsilon = \langle \cu \rangle$ (ony "almost" because not all combinations of partial transformations might exist in the paths). However $\cu \ltimes V$ associativy law doesn't have the limitation of~$\Upsilon$'s.
\end{remark}

\paragraph{(EC*) Path convolution operators}
The counterparts of strictly edge-constrained (EC*) convolution operators for path convolutions, are indeed path convolution operators obtained with bounded supporting set $\cn \subset \ct \ltimes_1 G$ which any graph can admit. As shown by this section, to construct one, all we need is a traversal set of partial transformations $(\cu, \ct)$.

